Foreground Signatures and Mitigation in Neutral Hydrogen Power Spectra From The Reionization Epoch

### 2014-10-19 foreground_delay_spectrum_modeling.tex
From 910390d: ============
(*) Made very minor change in abstract
(*) Addressed most of Gianni's comments
(*) Added few more references

\def\SKASA{\altaffilmark{2}}
\def\SKASAtxt{\altaffiltext{2}{Square Kilometre Array South Africa (SKA SA), Cape Town 7405, South Africa}}

\def\ANU{\altaffilmark{3}}
\def\ANUtxt{\altaffiltext{3}{Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia}}

\def\Haystack{\altaffilmark{4}}
\def\Haystacktxt{\altaffiltext{4}{MIT Haystack Observatory, Westford, MA 01886, USA}}

\def\Curtin{\altaffilmark{5}}
\def\Curtintxt{\altaffiltext{5}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}

@@

\def\MIT{\altaffilmark{6}}
\def\MITtxt{\altaffiltext{6}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}

\def\CfA{\altaffilmark{7}}
\def\CfAtxt{\altaffiltext{7}{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}}

\def\UW{\altaffilmark{8}}
\def\UWtxt{\altaffiltext{8}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}

\def\Victoria{\altaffilmark{9}}
\def\Victoriatxt{\altaffiltext{9}{School of Chemical \& Physical Sciences, Victoria University of Wellington, Wellington 6140, New Zealand}}

\def\UWisc{\altaffilmark{10}}
\def\UWisctxt{\altaffiltext{10}{Department of Physics, University of Wisconsin--Milwaukee, Milwaukee, WI 53201, USA}}

\def\UMichigan{\altaffilmark{11}}
\def\UMichigantxt{\altaffiltext{11}{Department of Atmospheric, Oceanic and Space Sciences, University of Michigan, Ann Arbor, MI 48109, USA}}

\def\CASS{\altaffilmark{12}}
\def\CASStxt{\altaffiltext{12}{CSIRO Astronomy and Space Science (CASS), PO Box 76, Epping, NSW 1710, Australia}}

\def\CAASTRO{\altaffilmark{13}}
\def\CAASTROtxt{\altaffiltext{13}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}}

\def\Tata{\altaffilmark{14}}
\def\Tatatxt{\altaffiltext{14}{National Centre for Radio Astrophysics, Tata Institute for Fundamental Research, Pune 411007, India}}

\def\RRI{\altaffilmark{15}}
\def\RRItxt{\altaffiltext{15}{Raman Research Institute, Bangalore 560080, India}}

\def\NRAO{\altaffilmark{16}}
\def\NRAOtxt{\altaffiltext{16}{National Radio Astronomy Observatory, Charlottesville and Greenbank, USA}}

\def\UMelbourne{\altaffilmark{17}}
\def\UMelbournetxt{\altaffiltext{17}{School of Physics, The University of Melbourne, Parkville, VIC 3010, Australia}}

@@

%% Author list
\author{
%% Lead Authors
Nithyanandan~Thyagarajan\ASU,
Daniel~C.~Jacobs\ASU,
Judd~D.~Bowman\ASU,
%% Supporting authors
% A.~P.~Beardsley\UW,
% J.~Pober\UW,
% I.~S.~Sullivan\UW,
%% Builders' list
G.~Bernardi\SKASA,
F.~Briggs\ANU,
R.~J.~Cappallo\Haystack, 
B.~E.~Corey\Haystack, 
% A.~A.~Deshpande\RRI, 
D.~Emrich\Curtin,
% B.~M.~Gaensler\USydney$^,$\CAASTRO, 
R.~Goeke\MIT,
L.~J.~Greenhill\CfA,
B.~J.~Hazelton\UW, 
M.~Johnston-Hollitt\Victoria,
D.~L.~Kaplan\UWisc, 
J.~C.~Kasper\UMichigan$^,$\CfA, 
E.~Kratzenberg\Haystack, 
C.~J.~Lonsdale\Haystack, 
M.~J.~Lynch\Curtin, 
S.~R.~McWhirter\Haystack,
D.~A.~Mitchell\CASS$^,$\CAASTRO, 
M.~F.~Morales\UW, 
E.~Morgan\MIT, 
D.~Oberoi\Tata, 
S.~M.~Ord\Curtin$^,$\CAASTRO,
T.~Prabu\RRI, 
A.~E.~E.~Rogers\Haystack, 
A.~Roshi\NRAO, 
N.~Udaya~Shankar\RRI, 
K.~S.~Srivani\RRI, 
R.~Subrahmanyan\RRI$^,$\CAASTRO, 
S.~J.~Tingay\Curtin$^,$\CAASTRO, 
M.~Waterson\Curtin$^,$\ANU,
R.~B.~Wayth\Curtin$^,$\CAASTRO, 
R.~L.~Webster\UMelbourne$^,$\CAASTRO, 
A.~R.~Whitney\Haystack, 
A.~Williams\Curtin, 
C.~L.~Williams\MIT,
and~MWA~EoR~members
}

%Institutional footnotes (typeset, then rearrange here to be in order)
\ASUtxt
\SKASAtxt
\ANUtxt
\Haystacktxt
\Curtintxt
% \USydneytxt
\MITtxt
\CfAtxt
\UWtxt
\Victoriatxt
\UWisctxt
\UMichigantxt
\CASStxt
\CAASTROtxt
\Tatatxt
\RRItxt
\NRAOtxt
\UMelbournetxt

@@

Detection of 21~cm emission of neutral hydrogen from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.  We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function.  Using the delay spectrum technique, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. Comparing data from recent Murchison Widefield Array observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings. For signals entering through the primary field of view, compact objects dominate the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space.  Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate foreground contamination substantially (up to a factor $\sim 10$ with negligible loss of data) in redshifted 21~cm power spectrum analysis.

@@

Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}. Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{tin13,bow13,lon09}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{thy13,bea13,bow06,pob14}.

A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}.

Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,liu14a,liu14b}. A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}). The more aggressive alternative is to subtract a sky model and regain access to modes lost by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here, we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.

Foregrounds that have intrinsic spectral variations, instruments with high chromaticity, or approximations in power spectrum analyses can contaminate the true EoR H{\sc i} power spectrum. Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity. To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual baseline--based approach of \citet{pob13} and \citet{par14}. This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.

Matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline. We report two important findings: the foregrounds that most severely obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a larger role than previously thought. % We find that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon. We find that diffuse Galactic emission plays a larger role then previously thought. 
We quantify this contamination by separating the simulations into components based on type of foreground emission. We then arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

@@

The observations shown here were undertaken on Aug~23, 2013. Data set from this day has been chosen as a reference for comparison within the MWA EoR collaboration because this is one of the days featuring a low number of known system errors from among a long stretch of stable observing.

Here we have chosen two, 112 second long sections from this night for detailed study. These were chosen to provide a selection of possible foreground and instrumental conditions. As an example of a nominal observing setup we choose a zenith pointing; as an example of poor foreground conditions, we choose a pointing when the field is $\sim 2$~hours from zenith. This pointing has a significantly higher secondary lobe structure and is observed when the bright galactic center is well above the horizon.

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al. 2014, submitted). MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning; it has a flux density limit of 25~mJy and a declination range of -12\arcdeg~ to -40\arcdeg~ evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern. The calibration algorithm -- based on forward modeling software by \citet{sul12} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay. Finally, we fit for an additional phase known to be caused by reflections in a few 150~m length cables. %possibly could add elaboration, if I knew exactly which version of the cal was done here. doesn't really matter

@@

The {\it off--zenith} pointing has emission notably higher than in the {\it zenith} pointing inside the {\it foreground wedge}. This is shown later to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky. Consequently, the contamination into the {\it EoR window} is also higher. Emission in the {\it zenith} pointing is more centrally concentrated due to phasing of power--pattern to the zenith, whereas the {\it off--zenith} pointing has a power--pattern phased eastward of zenith but is also responsive to emission from angles far away from zenith.

@@

Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the measurements perfectly, dominant at short antenna spacings where the number of delay modes within the horizon is lesser. This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum. However, we wish to emphasize that the purpose of deconvolution is to rid the delay spectra of instrumental artifacts to the best extent possible in order to see the underlying signatures of foreground emission. It is not our intention in this paper to use the deconvolution as a foreground removal tool.

@@

We describe the instrumental model and foreground models we use in our simulations.

@@

The instrument model consists primarily of the tile beam power--pattern, $A(\hat{s},f)$ (see equation~\ref{eqn:obsvis}). It is modeled as a mutually--coupled 4--by--4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14}. To speed up simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles. We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain \citep{bow07b}. Besides having the effect of adding a time--dependent uncertainty in the power pattern, these random delay fluctuations reduce the coherence in the phased addition of dipole signals resulting in deviations from predicted models of the power--pattern, most prominently at its nulls. The power--patterns so obtained are shown in Figure~\ref{fig:power_pattern} for the pointings chosen in our study.

@@

Figure~\ref{fig:data-sim-ratio} shows normalized histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}. The median absolute deviation from these distributions corresponds to $\sim$~90\% fractional difference between data and modeling on average in either case. This is in line with the aforementioned uncertainty in foreground models and thermal noise fluctuations in measurements. Currently we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
\caption{Normalized histograms of logarithm of ratios of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases. A value of zero indicates $\rho=1$. \label{fig:data-sim-ratio}}
\end{figure}

These uncertainties are presented to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}. These uncertainty estimates are conservative. A full treatment of all uncertainties and deviations from ideal behavior such as uncertainties in tile power--pattern, calibration, data corruption due to interference, and wide--field imaging effects will bring the simulations much closer in agreement with observations. Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.

@@

Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional delay spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}. The diffuse, compact, and composite components are are shown as dotted, dashed, and solid lines, respectively. The horizon delay limits are shown as a pair of vertical dot--dashed lines. The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum. Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the $\gtrsim$~70\% error owing to uncertainties in predicting the spectral index of compact foreground objects, while thermal noise fluctuations dominate outside.

@@

Near the horizon delay limits at shorter antenna spacing, diffuse emission is brighter than that from compact objects. Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at zero--delay. In fact, even on widely spaced antennas, {\emph diffuse emission in the delay spectrum near the horizon delay limits exceeds that in the main lobe by more than two orders of magnitude}. We attribute this to apparent shortening of antenna spacings projected in the direction of diffuse emission entering far from the primary field of view. This demonstrates the edge--heavy {\it two--pronged fork} feature discussed in \S\ref{sec:diffuse-long-baselines}.

Full simulations show the combination of center--heavy features on all antenna spacings from compact objects and edge--heavy features from both components especially diffuse emission even on wide antenna spacings.  This results in a characteristic {\it pitchfork} structure imprinted in the {\it foreground} wedge and should be evident in observations.

@@

Looking deeper at our simulations, below the level of the instrumental noise,  we identify numerous signatures of different components of foreground emission seen in the delay spectra. We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power--pattern and bandpass shape, and foreground parameters such as the nature of emission, spectral index, etc.

>>>>

\def\SKASA{\altaffilmark{2}}
\def\SKASAtxt{\altaffiltext{2}{Square Kilometre Array South Africa (SKA SA), Park Road, Pinelands 7405, South Africa}}

\def\RU{\altaffilmark{3}}
\def\RUtxt{\altaffiltext{3}{Department of Physics and Electronics, Rhodes University, Grahamstown 6140, South Africa}}

\def\CfA{\altaffilmark{4}}
\def\CfAtxt{\altaffiltext{4}{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}}

\def\ANU{\altaffilmark{5}}
\def\ANUtxt{\altaffiltext{5}{Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia}}

\def\Haystack{\altaffilmark{6}}
\def\Haystacktxt{\altaffiltext{6}{MIT Haystack Observatory, Westford, MA 01886, USA}}

\def\Curtin{\altaffilmark{7}}
\def\Curtintxt{\altaffiltext{7}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}

@@

\def\MIT{\altaffilmark{8}}
\def\MITtxt{\altaffiltext{8}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}

\def\UW{\altaffilmark{9}}
\def\UWtxt{\altaffiltext{9}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}

\def\Victoria{\altaffilmark{10}}
\def\Victoriatxt{\altaffiltext{10}{School of Chemical \& Physical Sciences, Victoria University of Wellington, Wellington 6140, New Zealand}}

\def\UWisc{\altaffilmark{11}}
\def\UWisctxt{\altaffiltext{11}{Department of Physics, University of Wisconsin--Milwaukee, Milwaukee, WI 53201, USA}}

\def\UMichigan{\altaffilmark{12}}
\def\UMichigantxt{\altaffiltext{12}{Department of Atmospheric, Oceanic and Space Sciences, University of Michigan, Ann Arbor, MI 48109, USA}}

\def\CASS{\altaffilmark{13}}
\def\CASStxt{\altaffiltext{13}{CSIRO Astronomy and Space Science (CASS), PO Box 76, Epping, NSW 1710, Australia}}

\def\CAASTRO{\altaffilmark{14}}
\def\CAASTROtxt{\altaffiltext{14}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}}

\def\Tata{\altaffilmark{15}}
\def\Tatatxt{\altaffiltext{15}{National Centre for Radio Astrophysics, Tata Institute for Fundamental Research, Pune 411007, India}}

\def\RRI{\altaffilmark{16}}
\def\RRItxt{\altaffiltext{16}{Raman Research Institute, Bangalore 560080, India}}

\def\NRAO{\altaffilmark{17}}
\def\NRAOtxt{\altaffiltext{17}{National Radio Astronomy Observatory, Charlottesville and Greenbank, USA}}

\def\UMelbourne{\altaffilmark{18}}
\def\UMelbournetxt{\altaffiltext{18}{School of Physics, The University of Melbourne, Parkville, VIC 3010, Australia}}

@@

%% Author list
\author{
%% Lead Authors
Nithyanandan~Thyagarajan\ASU,
Daniel~C.~Jacobs\ASU,
Judd~D.~Bowman\ASU,
%% Supporting authors
% A.~P.~Beardsley\UW,
% J.~Pober\UW,
% I.~S.~Sullivan\UW,
%% Builders' list
G.~Bernardi\SKASA$^,$\RU$^,$\CfA,
F.~Briggs\ANU,
R.~J.~Cappallo\Haystack, 
B.~E.~Corey\Haystack, 
% A.~A.~Deshpande\RRI, 
D.~Emrich\Curtin,
% B.~M.~Gaensler\USydney$^,$\CAASTRO, 
R.~Goeke\MIT,
L.~J.~Greenhill\CfA,
B.~J.~Hazelton\UW, 
M.~Johnston-Hollitt\Victoria,
D.~L.~Kaplan\UWisc, 
J.~C.~Kasper\UMichigan$^,$\CfA, 
E.~Kratzenberg\Haystack, 
C.~J.~Lonsdale\Haystack, 
M.~J.~Lynch\Curtin, 
S.~R.~McWhirter\Haystack,
D.~A.~Mitchell\CASS$^,$\CAASTRO, 
M.~F.~Morales\UW, 
E.~Morgan\MIT, 
D.~Oberoi\Tata, 
S.~M.~Ord\Curtin$^,$\CAASTRO,
T.~Prabu\RRI, 
A.~E.~E.~Rogers\Haystack, 
A.~Roshi\NRAO, 
N.~Udaya~Shankar\RRI, 
K.~S.~Srivani\RRI, 
R.~Subrahmanyan\RRI$^,$\CAASTRO, 
S.~J.~Tingay\Curtin$^,$\CAASTRO, 
M.~Waterson\Curtin$^,$\ANU,
R.~B.~Wayth\Curtin$^,$\CAASTRO, 
R.~L.~Webster\UMelbourne$^,$\CAASTRO, 
A.~R.~Whitney\Haystack, 
A.~Williams\Curtin, 
C.~L.~Williams\MIT,
and~MWA~EoR~members
}

%Institutional footnotes (typeset, then rearrange here to be in order)
\ASUtxt
\SKASAtxt
\RUtxt
\CfAtxt
\ANUtxt
\Haystacktxt
\Curtintxt
% \USydneytxt
\MITtxt
\UWtxt
\Victoriatxt
\UWisctxt
\UMichigantxt
\CASStxt
\CAASTROtxt
\Tatatxt
\RRItxt
\NRAOtxt
\UMelbournetxt

@@

Detection of 21~cm emission of neutral hydrogen from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.  We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function.  Using the delay spectrum technique, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. Comparing data from recent Murchison Widefield Array observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings. For signals entering through the primary field of view, compact objects dominate the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space.  Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate foreground contamination substantially (up to a factor $\sim 10$ with almost no loss of data) in redshifted 21~cm power spectrum analysis.

@@

Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}. Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{tin13,bow13,lon09}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,bea13,dil13,thy13,pob14}.

A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}.

Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,liu14a,liu14b}. A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}). The more aggressive alternative is to subtract a sky model and regain access to modes discarded by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here, we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.

Foregrounds with intrinsic deviations from spectral smoothness, instruments with high chromaticity, polarization leakage, calibration errors, or approximations in power spectrum analyses can contaminate the true EoR H{\sc i} power spectrum. Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity. To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual baseline--based approach of \citet{pob13} and \citet{par14}. This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.

After matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline. We report two important findings: the foregrounds that most severely obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. % We find that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon. We find that diffuse Galactic emission plays a larger role then previously thought. 
We quantify this contamination by separating the simulations into components based on type of foreground emission. We then arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

@@

The observations shown here were undertaken on Aug~23, 2013. % Data set from this day has been chosen as a reference for comparison within the MWA EoR collaboration because this is one of the days featuring a low number of known system errors from among a long stretch of stable observing.  
We have chosen two, 112 second long sections from this night for detailed study. These were chosen to provide a selection of possible foreground and instrumental conditions. As an example of a nominal observing setup we choose a zenith pointing; as an example of poor foreground conditions, we choose a pointing when the field is $\sim 2$~hours from zenith. This pointing has a significantly higher secondary lobe structure and is observed when the bright galactic center is well above the horizon.

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al. 2014, submitted). % MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning;
It has a flux density limit of 25~mJy and a declination range of -12\arcdeg~ to -40\arcdeg~ evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern. The calibration algorithm -- based on forward modeling software by \citet{sul12} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay. Finally, we fit for an additional phase known to be caused by reflections in a few 150~m length cables. %possibly could add elaboration, if I knew exactly which version of the cal was done here. doesn't really matter

@@

The {\it off--zenith} pointing has emission notably higher than in the {\it zenith} pointing inside the {\it foreground wedge}. This is shown later to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky. Consequently, the contamination into the {\it EoR window} is also higher, especially on short antenna spacings. Emission in the {\it zenith} pointing is more centrally concentrated due to phasing of power--pattern to the zenith, whereas the {\it off--zenith} pointing has a power--pattern phased eastward of zenith but is also responsive to emission from angles far away from zenith.

@@

% Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the measurements perfectly, dominant at short antenna spacings where the number of delay modes within the horizon is lesser. This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum. However, we wish to emphasize that the purpose of deconvolution is to rid the delay spectra of instrumental artifacts to the best extent possible in order to see the underlying signatures of foreground emission. It is not our intention in this paper to use the deconvolution as a foreground removal tool.

@@

We describe the instrumental and foreground models we use in our simulations.

@@

The instrument model consists primarily of the tile beam power--pattern, $A(\hat{s},f)$ (see equation~\ref{eqn:obsvis}). It is modeled as a mutually--coupled 4--by--4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14}. To speed up simulations, we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles. We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain \citep{bow07b}. Besides having the effect of adding a time--dependent uncertainty in the power pattern, these random delay fluctuations reduce the coherence in the phased addition of dipole signals resulting in deviations from predicted models of the power--pattern, most prominently at its nulls. The power--patterns so obtained are shown in Figure~\ref{fig:power_pattern} for the pointings chosen in our study.

@@

Figure~\ref{fig:data-sim-ratio} shows normalized histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}. The median absolute deviation from these distributions corresponds to $\sim$~90\% fractional difference between data and modeling on average in either case. This is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account. Currently, we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
\caption{Normalized histograms of logarithm of ratios of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases indicating a fractional difference of $\sim 90$\%. A value of zero indicates perfect agreement ($\rho=1$). \label{fig:data-sim-ratio}}
\end{figure}

These uncertainties are presented to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}. These uncertainty estimates are conservative. A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power--pattern \citep{ber14}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations. Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.

@@

Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional delay spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}. The diffuse, compact, and composite components are shown as dotted, dashed, and solid lines, respectively. The horizon delay limits are shown as a pair of vertical dot--dashed lines. The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum. Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the $\gtrsim$~70\% error owing to uncertainties in predicting the spectral index of compact foreground objects, while thermal noise fluctuations dominate outside.

@@

Near the horizon delay limits at shorter antenna spacing, diffuse emission is brighter than that from compact objects. Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at zero--delay. In fact, even on widely spaced antennas, {\emph diffuse emission in the delay spectrum near the horizon delay limits exceeds that in the main lobe by more than two orders of magnitude}. We attribute this to apparent shortening of antenna spacings projected in the direction of diffuse emission entering far from the primary field of view, thereby giving rise to the edge--heavy {\it two--pronged fork} feature discussed in \S\ref{sec:diffuse-long-baselines}.

Simulations with the complete foreground model show the combination of center--heavy features on all antenna spacings from compact emission, and edge--heavy features from both compact and diffuse emission especially the latter even on wide antenna spacings.  This results in a characteristic {\it pitchfork} structure imprinted in the {\it foreground} wedge and should be evident in observations.

@@

Our simulations enable us to identify numerous signatures of different components of foreground emission seen in the delay spectra. We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power--pattern, and foreground parameters such as the nature of emission, spectral index, etc.
